---
title: "p8105_hw1_tk2886"
author: "Tanvir Khan"
date: 2021-09-24
output: github_document
---


```{r setup, include=FALSE}
library(tidyverse)
```

## QUESTION 1
```{r Creating a dataframe}
set.seed(12345)

mydataframe <- tibble(
                  random_sample = rnorm(10),
                  logical_vec = c(random_sample > 0),
                  char_vec = c("My", "name", "is", "Tanvir", "Khan", "and", "I", "like","the", "Lakers"),
                  fac_vec = factor(c("single", "single", "single", "single", "married", "married", "married", "divorced", "divorced", "divorced")),
)

knitr::kable(mydataframe)
```


```{r Mean Analysis, collapse = TRUE}
rnorm_mean = mean(pull(mydataframe, random_sample))
logicalvec_mean = mean(pull(mydataframe, logical_vec))
charvec_mean = mean(pull(mydataframe, char_vec))
facvec_mean = mean(pull(mydataframe, fac_vec))
```
When I take the mean of the random sample of size 10 variable from a standard Normal distribution, the mean function works on this and it gives me a value. This works because it contains numeric value. I can take the mean of a numeric value. 

When I take the mean of the logical vector, the mean functions work on this logical vector even though that does not make sense. If someone looks at the data frame they will see that the logical vector in my dataframe displays True and False. The True is being converted to 1 and False is being converted to 0. I have four True out of 10 and that is 0.4 which is the same value as the mean function is outputting. 

When I take the mean of the character vector and factor vector, the console output is displaying a warning message that states the argument is not numeric or logical: "Returning NA". This makes sense for the character vector since I can not take the mean of a character vector which does not have numeric values. The factor vector is not numeric or logical. 


```{r convert to numeric, collapse = TRUE}
numeric_char = as.numeric(pull(mydataframe,char_vec))
numeric_logv = as.numeric(pull(mydataframe, logical_vec))
numeric_fac = as.numeric(pull(mydataframe, fac_vec))
```
When I attempted to convert the character variables in the character vector, NA is returned. If I understand correctly, the values in the character value are characters/letters and we can't convert words or letters to numbers. When I convert the logical variables from the logical vector to numeric, it displays 0 and 1. 0's indicate True and 1's indicate False. When I convert factor variables in the factor vector to numeric, 3,2,1 are displayed. I believe that the 3,2,1 are numeric codes for the three different levels of the factor. 


```{r reattempting to convert character to numeric}
numeric_character = as.numeric(factor(pull(mydataframe, char_vec)))
```




```{r calculating the mean of the new conversion numbers}
mean_numchar = mean(numeric_char)
mean_numlogic = mean(numeric_logv)
mean_numfactor = mean(numeric_fac)
```
When I take the mean of the converted character vector, it gives me NA because I did not successfully convert character vector to numeric variables. When I take the mean of the numeric converted logical vector, I get the mean of 0.4 because R is handling True's as 0 and False as 1. When I take the mean of the numeric converted factor vector, I get the mean of 2.1 which is the mean of the different levels of the factor vector. 


## QUESTION 2
```{r}
data("penguins", package = "palmerpenguins")
skimr::skim(penguins)
```
a) The important variables are *`r names(penguins)`*.
b) The unique **species** in the penguins data table are *`r unique(penguins$species)`*.
c) The unique **island** in the penguins data table are *`r unique(penguins$island)`*.
d) The unique **years** in the penguins data table are *`r unique(penguins$year, na.rm = TRUE)`*.
e) The **sex** in the penguins data table are *`r unique(na.omit(penguins$sex))`*.
e) The mean of the **bill length** is *`r round(mean(penguins$bill_length_mm, na.rm = TRUE), digits = 2)`* and the standard deviation of the bill length is *`r round(sd(penguins$bill_length_mm, na.rm = TRUE), digits = 2)`*.
f) The mean of the **bill depth** is *`r round(mean(penguins$bill_depth_mm, na.rm = TRUE), digits = 2)`* and the standard deviation of the bill depth is *`r round(sd(penguins$bill_depth_mm, na.rm = TRUE), digits=2)`*.
g) The mean of the **body mass** is *`r round(mean(penguins$body_mass_g, na.rm = TRUE), digits=2)`* and the standard deviation of the body mass is *`r round(sd(penguins$body_mass_g, na.rm = TRUE), digits=2)`*.
h) There is *`r nrow(penguins)`* **rows** in this data table. There is *`r ncol(penguins)`* **columns** in this data table. 
i) The mean of the **flipper length** is *`r round(mean(penguins$flipper_length_mm, na.rm = TRUE))`* and the standard deviation of the flipper lenght is *`r round(sd(penguins$flipper_length_mm, na.rm = TRUE))`*. 

```{r Creating scatterplot}
penguin_scatterplot <- ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point(size = 2, shape = 23)

penguin_scatterplot
```

```{r Saving ggplot as PDF}
ggsave("scatter_plot.pdf", plot = penguin_scatterplot, height = 4, width = 6)
```


