---
title: "p8105_hw1_tk2886"
author: "Tanvir Khan"
date: 2021-09-24
output: github_document
---


```{r setup, include = FALSE}
library(tidyverse)
```



# QUESTION 1 - Creating a dataframe 

```{r Creating a dataframe}

set.seed(12345)

mydataframe <- tibble(
                  random_sample = rnorm(10),
                  logical_vec = random_sample > 0,
                  char_vec = c("My", "name", "is", "Tanvir", "Khan", "and", "I", "like","the", "Lakers"),
                  fac_vec = factor(c("single", "single", "single", "single", "married", "married", "married", "divorced", "divorced", "divorced")),
)

knitr::kable(mydataframe)
```




## Calculating the mean of each of the variable in my dataframe 

```{r Mean Analysis, collapse = TRUE}
rnorm_mean = mean(pull(mydataframe, random_sample))
logicalvec_mean = mean(pull(mydataframe, logical_vec))
charvec_mean = mean(pull(mydataframe, char_vec))
facvec_mean = mean(pull(mydataframe, fac_vec))

rnorm_mean
logicalvec_mean
charvec_mean
facvec_mean
```
When I take the mean of the random sample of size 10 variable from a standard Normal distribution, the mean function works on this and it gives me a value. This works because it contains numeric value. I may take the mean of a numeric value. 

When I take the mean of the logical vector, the mean functions work on this logical vector even though that does not make sense. If someone looks at the data frame they will see that the logical vector in my data frame displays True and False. I think the True is being converted to 1 and False is being converted to 0. I have four True out of 10 and that is 0.4 which is the same value as the mean function is outputting. 

When I take the mean of the character vector and factor vector, the console output is displaying a warning message that states the argument is not numeric or logical: "Returning NA". This makes sense for the character vector since I can not take the mean of a character vector which does not have numeric values. The factor vector is not numeric or logical. 




## 1st attempt in converting character, logcial, and factor variables into numeric 

```{r 1st attempt to convert to numeric, message = FALSE, warning = FALSE}
numeric_char = as.numeric(pull(mydataframe,char_vec))
numeric_logv = as.numeric(pull(mydataframe, logical_vec))
numeric_fac = as.numeric(pull(mydataframe, fac_vec))
```

When I attempted to convert the elements in the character vector, NA is returned. The elements in the character vector are characters/letters and we can not convert words or letters to numbers. When I convert the elements from the logical vector to numeric, it displays 0 and 1. 1's indicate True and 0's indicate False. When I convert elements of the factor vector to numeric, 3,2,1 are displayed. I believe that the 3,2,1 are numeric codes for the three different levels of the factor. 




## Calculating the mean of the newly converted numeric variables 

```{r calculating the mean of the new conversion numbers, message = FALSE, warning = FALSE}
mean_numchar = mean(numeric_char)
mean_numlogic = mean(numeric_logv)
mean_numfactor = mean(numeric_fac)
```

When I take the mean of the converted character variable, it gives me NA because I did not successfully convert character variable to numeric variable. When I take the mean of the numeric converted logical variable, I get the mean of 0.4 because R is handling True's as 0 and False as 1. When I take the mean of the numeric converted factor vector, I get the mean of 2.1 which is the mean of the different levels of the factor vector. 



## 2nd attempt in converting factor variable to numeric 

```{r 2nd attempt at converting factor variable to numeric, message = FALSE, warning = FALSE}
numeric_factorvec = as.numeric(as.character(pull(mydataframe, fac_vec)))
```

After doing some research on Google, I found out that since my factor vector has character text like "single", "married", "divorce", I will not be able to convert to numeric even if I use the as.character() function. If the factor vector contained elements that are numbers like c(123,243,3232,412,51) then I may convert that to character by using as.character(). This will represent those exact numbers in text format and then I may do one additional conversion to numeric by using the as.numeric(). This will give me the exact values that was stored in the factor vector. 




# QUESTION 2 - Practicing with inline R code, plotting, and behavior of ggplot

```{r}
data("penguins", package = "palmerpenguins")
```



### Practicing with inline R code and extracting specific data from penguins data

a) The important variables are *`r names(penguins)`*.

b) The number of **rows** in the penguins data set is **`r nrow(penguins)`**.

c) The number of **columns** in the penguins data set is **`r ncol(penguins)`**.

d) The unique **species** in the penguins data set are **`r unique(pull(penguins, species))`**.

e) The unique **island** in the penguins data set are **`r unique(pull(penguins, island))`**.

f) The unique **years** in the penguins data set are **`r unique(pull(penguins, year), na.rm = TRUE)`**.

g) The **sex** in the penguins data set are **`r unique(na.omit(pull(penguins, sex)))`**.

h) 
The mean of the **bill length** is *`r round(mean(pull(penguins, bill_length_mm), na.rm = TRUE), digits = 2)`*mm.
The standard deviation of the bill length is *`r round(sd(pull(penguins, bill_length_mm), na.rm = TRUE), digits = 2)`*.

i) 
The mean of the **bill depth** is *`r round(mean(pull(penguins, bill_depth_mm), na.rm = TRUE), digits = 2)`*mm.
The standard deviation of the bill depth is *`r round(sd(pull(penguins, bill_depth_mm), na.rm = TRUE), digits=2)`*.

j) 
The mean of the **flipper length** is *`r round(mean(pull(penguins, flipper_length_mm),na.rm = TRUE))`*mm.
The standard deviation of the flipper length is  *`r round(sd(pull(penguins, flipper_length_mm), na.rm = TRUE), digits=2)`*.

k)
The mean of the **body mass** is *`r round(mean(pull(penguins, body_mass_g), na.rm = TRUE), digits=2)`*mm.
The standard deviation of the body mass is *`r round(sd(pull(penguins, body_mass_g), na.rm = TRUE), digits=2)`*.




### First time creating scatterplot for this class 
```{r Creating scatterplot, message = FALSE, warning = FALSE}
penguin_scatterplot <- ggplot(penguins, aes(x = bill_length_mm, y = flipper_length_mm, color = species)) +
  geom_point(shape = penguins$species, size = 3) +
  ggtitle("Flipper Length vs Bill Length") + theme(plot.title = element_text(hjust = 0.5))

penguin_scatterplot
```


### Saving my first ggplot (scatterplot) as PDF
```{r Saving ggplot as PDF, message = FALSE, warning = FALSE}
ggsave("scatter_plot.pdf", plot = penguin_scatterplot, height = 4, width = 6)
```


